## Peer Assessment Practical Machine Learning
#### Predict activity type

## Load the data
```{r cache=TRUE}
set.seed(1234)
library(caret)
location <- '/home/wijnand/R_workspace/8_Practical_Machine_Learning/course_project/resources/train.csv'
data <- read.csv(location, na.strings = c("NA", "#DIV/0!", ""))
```
## Modify data
#### Remove 'zero' and 'near-zero' columns
```{r cache=TRUE}
data$X <- NULL
uselessColumns <- nearZeroVar(data, freqCut = 95/5, uniqueCut = 10, saveMetrics = T)
uselessColumns <- names(data[,uselessColumns$zeroVar==TRUE | uselessColumns$nzv==TRUE])
data <- data[,!(names(data) %in% uselessColumns)]
```     

####  Remove columns in which > 97% of the values are not available
```{r cache=TRUE}
data <- data[,colSums(is.na(data))<(nrow(data)*0.97)]
```

## Prepare data
#### Randomize the dataset
```{r cache=TRUE}
data <- data[order(runif(nrow(data))),]
```
#### Create training & test part
```{r cache=TRUE}
inTrain <- createDataPartition(data$classe, p = 0.7, list=F)
training <- data[inTrain,]
testing <- data[-inTrain,]
```

## Train a random forest model
```{r cache=TRUE}
trainedModel <- train(classe ~ . , training, method="rf", trControl = trainControl(method = "cv", number = 2),do.trace=F,ntree=50)
```
## Predict
#### Predict the test data using the created random-forest-model
```{r pred1, cache=TRUE}
prediction <- predict(trainedModel, testing)
```

#### Print the results
```{r cache=TRUE}
confusionMatrix(prediction, testing$classe)
```
The prediction accuracy is .998 on the 30% test-data which is awesome! 

#### Predict unseen testdata
```{r cache=TRUE}
testdata <- read.csv('/home/wijnand/R_workspace/8_Practical_Machine_Learning/course_project/resources/test.csv', 
                             na.strings = c("NA", "#DIV/0!", ""))
prediction <- predict(trainedModel, testdata)
n <- length(prediction)
for(i in 1:n)
{
        filename <- paste0("/home/wijnand/R_workspace/8_Practical_Machine_Learning/course_project/resources/problem_id_",i,".txt")
        write.table(prediction[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
}
```
All 20 (unseen) cases were predicted correctly.
